Fast gravitational wave parameter estimation with no-compromise

### 2023-01-22 src/tex/ms.tex
turned comment into edit
removed undevided acr BBH and BNS

To compute the integral shown in Eq.~\eqref{eq: innerproduct}, we need to
evaluate a chosen waveform model $h(\mathbf{\theta})$ at a number of frequency
sample points. This makes evaluating the likelihood function often the most
computationally intensive part of PE. The most accurate waveform model is
numerical relativity (NR), which obtains waveforms by directly solving the
Einstein equations numerically for a given system. However, depending on the
source parameters, generating one time series of strain can take a day to half a
year, which makes NR prohibitively expensive for PE. To circumvent this problem,
there are several families of waveform ``approximants'', including the
\textsc{IMRPhenom} family \cite{Khan:2015jqa, Garcia-Quiros:2020qpx}, the
\textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$
BBH, one waveform call could take $10\text{ms}$ to ${\sim}1s$ \kw{verify number
in lalsuite}. For longer events, such as a $1.4-1.4\ M_{\odot}$ BNS, the
evaluation time could go up to \kw{Fill}. Since one needs to evaluate the
likelihood millions of times during sampling,\footnote{A typical PE run with
\texttt{Bilby} takes ${>}10^6$ likelihood evaluations to converge.} the
computational cost in evaluating the waveform accumulates and is the main reason
of the long runtime of GW PE.

@@

To obtain a reference waveform, we currently use the \texttt{differential
evolution} algorithm \cite{Storn1997DifferentialE} available in the
\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters which maximize the
likelihood.
\mi{The reference waveform could also be produced from trigger parameters
precomputed by a search pipeline without additional computation.}

>>>>

To compute the integral shown in Eq.~\eqref{eq: innerproduct}, we need to
evaluate a chosen waveform model $h(\mathbf{\theta})$ at a number of frequency
sample points. This makes evaluating the likelihood function often the most
computationally intensive part of PE. The most accurate waveform model is
numerical relativity (NR), which obtains waveforms by directly solving the
Einstein equations numerically for a given system. However, depending on the
source parameters, generating one time series of strain can take a day to half
a year, which makes NR prohibitively expensive for PE. To circumvent this
problem, there are several families of waveform ``approximants'', including the
\textsc{IMRPhenom} family \cite{Khan:2015jqa, Garcia-Quiros:2020qpx}, the
\textsc{SEOB} family \cite{PhysRevD.89.061502}, and the NR surrogate
family\cite{Varma:2019csw}. For shorter events, such as a $30-30\ M_{\odot}$
binary black hole, one waveform call could take $10\text{ms}$ to ${\sim}1s$
\kw{verify number in lalsuite}. For longer events, such as a $1.4-1.4\
M_{\odot}$ binary neutron star, the evaluation time could go up to \kw{Fill}.
Since one needs to evaluate the likelihood millions of times during
sampling,\footnote{A typical PE run with \texttt{Bilby} takes ${>}10^6$
likelihood evaluations to converge.} the computational cost in evaluating the
waveform accumulates and is the main reason of the long runtime of GW PE.

@@

To obtain a reference waveform, we currently use the \texttt{differential
evolution} algorithm \cite{Storn1997DifferentialE} available in the
\textsc{scipy} package \cite{2020SciPy-NMeth} to find the waveform parameters
which maximize the likelihood.  The reference waveform could also be produced
from trigger parameters precomputed by a search pipeline without additional
computation.
